Template Master's Thesis

### 2023-08-19 paper.tex
Update discussion

%The results chapter or section simply and objectively reports what you found, without speculating on why you found these results.

%The discussion
%... interprets the meaning of the results, puts them in context, and explains why they matter.
%... is an in-depth exploration of the results, going into detail about the meaning of your findings and citing relevant sources to put them in context.
%... Try to explain "What do your results mean?", and how they relate to the literature,
%... provides the explanation and interpretation of results or findings by comparing with the findings in prior studies.

Throughout our analysis, we successfully applied three heuristics to cluster our address set, while utilizing on-chain data. In the following discussion, we will examine each of the approaches step-by-step. %The heuristics we applied were self-authorization, node embedding, and LAND transfers.

%%%%%%%%%%%%%%%%%%%%
Implementing the self-authorization heuristic is straightforward, but results in very few identified address clusters. This is because it requires a highly specific on-chain behavior, which is rarely observed in practice. In particular, approvals are seldom used to authorize another externally owned account one might own. We also encountered a significant number of false-positives due to non-standard implementations. After excluding these instances, we identified a total of 15 entities consisting of 35 addresses. This outcome aligns with findings by \cite{FV:17}, who also observed that the heuristic captures only a small proportion of addresses. In his analysis of Ethereum transaction data up to block number 8,500,000 he identified 4,599 entities from 7,107 addresses.
%%%%%%%%%%%%%%%%%%%%

Address clustering through node embedding methods proved to be the most effective approach in our analysis. Unlike methods that require participation in specific on-chain events, node embedding examines the entire interaction network between addresses, making it broadly applicable. \newline
Our implementation of node embedding methods differs from previous approaches by \cite{Beres2020} and \cite{wu2022tutela}, the main reason being different settings and research questions. We construct the network graph based on the asset transfers within our address set and simultaneously analyze Polygon and Ethereum data. \newline
Conversely, \cite{wu2022tutela} aim to cluster addresses without a predefined set, therefore using transaction data for network graph construction. \cite{Beres2020} try to match deposit and withdrawal addresses for Tornado Cash, relying on all token transfers and transactions to cluster addresses based on similar on-chain behavior. Since it is not very common that Tornado cash withdrawal and deposit addresses interact with each other, the approach chosen by \cite{Beres2020} is better suited for their objectives.

Moreover, we validated our approach and found better results in terms of average rank of ENS pairs, when building the network graph out of intra-set asset transfers compared to using all token transfer events and/or transactions. \newline
%Moreover, when evaluating different network graphs as the starting point to create node embeddings we found better results in terms of average rank with the intra-set asset transfer dataset, instead of using all transfer events and/or transactions.
We applied three node embedding algorithms that showed strong performance in the experiments by \cite{Beres2020} and assessed their effectiveness. Our results indicate that Role2Vec is the best-performing method, locating the ground truth pair within the five nearest neighbors in 70\% of cases.  This diverges from previous findings that favored Diff2Vec as the most effective of the three methods. \newline
Determining why one approach outperforms another is complex. Diff2Vec and DeepWalk, both neighborhood-preserving algorithms, mainly capture node proximity. However, as noted by ..., this does not necessarily imply similarity, especially when addresses may frequently interact without being controlled by the same entity. In contrast, Role2Vec, a structural node embedding method, identifies nodes with similar network roles. In our scenario, network position or role may better represent similarity and structural relatedness among nodes in our evaluation sample. Despite these insights, the differences in results could be specific to the data at hand, given the small sample size, potential evaluation metric flaws, and the overall robust performance of all methods. \newline
Next, we created clusters from our node embeddings, enabling us to estimate the number of entities represented by the 51,566 addresses. Our method involves calculating the Euclidean distance between pairs of feature vectors, specifically using the node embedding matrix generated by Role2Vec. Nodes are clustered if their distance falls below a predetermined threshold, and any overlapping clusters are subsequently merged. \newline
Though our method may not be the ideal approach for clustering, it is likely more effective than simply clustering a fixed number of $k$ addresses. We observed that the number of clusters and their size is highly sensitive to the chosen distance threshold, making it challenging to assess the accuracy of these clusters. We have provided a sensitivity analysis to illustrate this issue. The distances between the feature vectors tend to be quite narrow (dense vectors), and even the distances between various ground truth pairs exhibit significant variation. \newline
We explored potential enhancements to our method, such as incorporating additional information, e.g., time-of-day-activity and/or gas price selection into the feature vectors. Unfortunately, these modifications did not yield improvements. Similarly, a two-step approach to clustering also proved unsuccessful. These areas remain open for future research.

%Clusters aber sehr abhängig von threshold oder der methode wie man sie bildet, möglichkeit wäre weitere information (time-of-day) mit einfliessen zu lassen (evtl. sogar in den vector einbauen) -> wir haben es getestet aber kein (signifikantes) improvement für unsere address pairs gefunden. Könnte aber auch sein dass wir es falsch angewendet haben.
%und Beres hat ein kleineres Set an addressen (er will nur deposit und withdraw addressen von tornado cash clustern) und benutzt all token transfers und transaktionen, nicht nur diese zwischen den addressen, was bei ihm auch nicht viele wären da er es mit tornado anschaut und es nicht sinnvoll wäre zwischen den addressen zu interagieren.
%Im Gegensatz zu früheren implementierungen haben wir intra-set asset transfers benutzt um den graph zu bilden. das ist so weil bei unser spezifischen aufgabe ein vordefiniertes grosses addressen set zu clustern dies besser gelingt als bspw. alle transaktionen im ethereum netzwerk anzuschauen wie tutela oder wie beres der einfach alle token transfers der withdrawal und deposit addressen von tornado cash anschaut, wo sie eher weniger interaktionen haben als bei uns.
%Node embeddings mussten wir an unsere Gegebenheiten anpassen. Evaluierung mittels unsern ENS address paaren, hat gezeigt dass diese methode sehr gut funktioniert um ähnliche addressen zu finden.
%Despite the findings of \cite{Beres2020}, it appears that Diff2Vec is slightly less effective than Role2Vec or DeepWalk, although it is still generally successful. 
%Schwierig zu sagen wieso der eine Appraoch besser ist als der andere, der unterschied zwischen r2v und den anderen zwei ist das r2v ein role-based algorithmus ist und dies für unser spezielles datenset und unsere ENS address pairs besser zu funktionieren scheint. Aber nicht sehr aussagekräftig da kleines sample und alle approaches recht gut.
%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%
-Eine weitere Methode basierend auf spezifischen on-chain Event vorgeschlagen.
-The large cluster may incorrectly be marked/maybe off-chain compensation? -> auction address
-lässt sich gut visualisieren und evtl. auch nachvollziehen
-eigener approach hat ein spezifisches on-chain event zur voraussetzung, basiert auf annahme dass LAND nicht einfach so weitergegeben wird, ist auch auf andere nfts oder grosse token transfers anwendbar. hat auch funktioniert aber wieder relativ wenige clusters gefunden, hat es falsch positive dabei? zb. offchain zahlungen und on-chain token transferiert
%%%%%%%%%%%%%%%%%%%%%

-Discuss which approach worked best for clustering a predefined address set and why. 
Diskutiere alle resultate gemeinsam, vom anfangs set über on-chain und intra-set zu den clusters.

>>>>

%The results chapter or section simply and objectively reports what you found, without speculating on why you found these results. 
%The discussion
%... interprets the meaning of the results, puts them in context, and explains why they matter.
%... is an in-depth exploration of the results, going into detail about the meaning of your findings and citing relevant sources to put them in context.
%... Try to explain "What do your results mean?", and how they relate to the literature,
%... provides the explanation and interpretation of results or findings by comparing with the findings in prior studies.

Throughout our analysis, we successfully applied three heuristics to cluster our address set while making use of on-chain data. In the following discussion, we will examine each of the approaches step-by-step.

%\textbf{Self-authorization}\\
Implementing the \textit{self-authorization} heuristic is straightforward, but results in very few identified address clusters. This is because it requires a highly specific on-chain behavior, which is rarely observed in practice. In particular, approvals are seldom used to authorize another externally owned account one might own. We also encountered a significant number of false-positives due to non-standard implementations. After excluding these instances, we identified a total of 15 entities consisting of 35 addresses. This outcome aligns with findings by \cite{FV:17}, who also observed that the heuristic captures only a small proportion of addresses. In his analysis of Ethereum transaction data up to block number 8,500,000 he identified 4,599 entities from 7,107 addresses.

%\textbf{Graph representation learning}\\
Address clustering through \textit{graph representation learning} proved to be the most effective approach in our analysis. Unlike methods that require participation in specific on-chain events, node embedding algorithms examine the entire interaction network between addresses, making it broadly applicable. \newline
Our implementation of node embedding methods differs from previous approaches by \cite{Beres2020} and \cite{wu2022tutela}, the main reason being different settings and research questions. We construct the network graph based on the asset transfers within our address set and simultaneously analyze Polygon and Ethereum data. \newline
Conversely, \cite{wu2022tutela} aim to cluster addresses without a predefined set, therefore using Ethereum transaction data for network graph construction. \cite{Beres2020} try to match deposit and withdrawal addresses for Tornado Cash, relying on all token transfers and transactions to cluster addresses based on similar on-chain behavior. Since it is not very common that Tornado cash withdrawal and deposit addresses interact with each other, the approach chosen by \cite{Beres2020} is better suited for their objectives. \newline
Moreover, we validated our approach and found better results in terms of average rank of ENS pairs, when building the network graph out of intra-set asset transfers compared to using all token transfer events and/or transactions. A possible extension would include incorporating edge weights, similar to \cite{wu2022tutela}, however, we refrained from doing this because we would have needed to implement a custom node embedding algorithm.\newline
We applied three node embedding algorithms that delivered robust results in the experiments by \cite{Beres2020} and assessed their effectiveness in detecting ground truth pairs. Our results indicate that Role2Vec is the best-performing method, locating the ground truth pair within the five nearest neighbors in 70\% of cases.  This diverges from the findings of \cite{Beres2020}, who determined Diff2Vec to be the superior method among the three evaluated. \newline
Determining why one approach outperforms another is challenging. DeepWalk and Diff2Vec, both neighborhood-preserving algorithms, mainly capture node proximity. As noted by \cite{ahmed2018roletovec}, this does not necessarily imply similarity, especially when addresses frequently interact with each other without being controlled by the same entity. In contrast, Role2Vec, a structural node embedding method, identifies nodes with similar roles within the network. In our evaluation sample, network position or role may better represent similarity and structural relatedness among nodes. Despite these insights, the differences in results could be specific to the data at hand, given the small sample size, potential evaluation metric flaws, and the overall robust performance of all methods. \newline
Creating clusters out of the node embeddings, enables us to estimate the number of entities represented by the 51,566 addresses. Our method involves calculating the Euclidean distance between pairs of feature vectors, specifically using the node embedding matrix generated by Role2Vec. Nodes are clustered if their distance falls below a predetermined threshold, and any overlapping clusters are subsequently merged. \newline 
While our clustering method may not be ideal, it is probably more effective than, for example, clustering a fixed number of $k$ neighbors. We observed that the number of clusters and their size is highly sensitive to the chosen distance threshold, making it challenging to assess the accuracy of these clusters. We have provided a sensitivity analysis to illustrate this issue. The distances between the feature vectors tend to be quite narrow, with significant variation even among various ground truth pairs' distances. Regardless of the distance threshold, we were able to locate a cluster of 500+ addresses. Manual inspection revealed that these address show very similar on-chain patterns: They are only active on Polygon, most of the transfers are NFT mints and all NFTs are somehow related to Decentraland. Further, they transfer specific NFTs to the same address. Unsere Vermutung ist dass es ein Bot ist der mit vielen verschiedenen Adressen auf Decentraland aktiv ist, an verschiedene Orte/Events geht um dort NFT airdrops zu erhalten. Wenn sie etwas wert sind aggregiert er die NFTs auf eine addresse und verkauft sie über den marketplace. Man könnte diese Vermutung mit den bewegungsdaten validieren/überprüfen. Spannend dass der node embedding algo trotz limiterten informationen dieses pattern entdeckt hat. Etwas das in zukunft ein problem werden könnte -> evtl. sehr viel Bot aktivität auf DCL und evtl. mehr Research reinfliessen könnte.

We explored potential enhancements to our method, such as incorporating additional information, e.g., time-of-day-activity and/or gas price selection into the feature vectors. Unfortunately, these modifications did not yield improvements. %vielleicht auch falsch gemacht
Similarly, a two-step approach to clustering also proved unsuccessful. These areas remain open for future research.

%Moreover, when evaluating different network graphs as the starting point to create node embeddings we found better results in terms of average rank with the intra-set asset transfer dataset, instead of using all transfer events and/or transactions.

%\textbf{LAND transfers}\\
The \textit{LAND transfer} heuristic again looks at very specific on-chain behavior.  740 transfers, 413 addressen und 161 or 160 entities with 342 addresses. -The large cluster may incorrectly be marked/maybe off-chain compensation? -> auction address

 We use the fact that these addresses are related to Decentraland and therefore choose LAND. But this could also be applied to other NFTs or large ERC20 token/Ether/stablecoin transfers. Main reasoning: revealing clusters by sending high-value amount to other address why? reorganize wallet(s), upgrade security, increased value of NFTs/assets over time -> change from metamask (starting option) to hardware wallet.

 Pro: lässt sich gut visualisieren und evtl. auch nachvollziehen
 Kontra: Likely a lot of false-positives with this appraoch, could be something missing in the on-chain data.

%eigener approach hat ein spezifisches on-chain event zur voraussetzung, basiert auf annahme dass LAND nicht einfach so weitergegeben wird, ist auch auf andere nfts oder grosse token transfers anwendbar. hat auch funktioniert aber wieder relativ wenige clusters gefunden, hat es falsch positive dabei? zb. offchain zahlungen und on-chain token transferiert
%%%%%%%%%%%%%%%%%%%%%

Diskutiere alle resultate gemeinsam, vom anfangs set über on-chain und intra-set zu den clusters.
470k -> 135k on chain -> 51k intra-set transfer -> etc.
